---
title: "Project 3 - Main Script"
author: "Alek Anichowski, Sophie Beiers, Mingyue Kong, Yun Li, Keith Rodriguez"
date: "March 5th, 2018"
output:
  pdf_document: default
  html_document: default
---
In your final Project 2 repo, there should be an R markdown file called `main.Rmd` that organizes **all computational steps** for evaluating your proposed image classification framework. 

This file is meant to be a template for evaluating models used for image analysis (and could be generalized for any predictive modeling). You should update it according to your models/codes but your final document should have precisely the same structure. 

```{r, warning = FALSE}
if(!require("EBImage")){
  source("https://bioconductor.org/biocLite.R")
  biocLite("EBImage")
}

if(!require("gbm")){
  install.packages("gbm")
}

if(!require("pbapply")){
  install.packages("pbapply")
}

if(!require("pbapply")){
  install.packages("caret")
}

if(!require("FNN")){
  install.packages("FNN")
}

library(FNN)
library("EBImage")
library("gbm")
library("pbapply")
library("caret")
library(randomForest)
```

## Step 0: Specify directories.

We first set the working directory to the location of this .Rmd file. Then we specify our training and testing data. To train the initial baseline model, we used the `caret` package to partition the data into training and testing sets. This code can be seen commented out. 

```{r wkdir, eval=FALSE}
setwd("~/Documents/GitHub/project-2-predictive-modelling-group-3/doc")
```

Now we provide directories for the raw images. Here we assume the training set and test set are in different subfolders. 
```{r}
experiment_dir <- "../data/pets/" 
img_train_dir  <- paste(experiment_dir, "train/", sep="")
img_test_dir   <- paste(experiment_dir, "test/", sep="") 
```

## Step 1: Set up controls for model evaluation.

In this step, we have a set of controls for the model evaluation.  The code in the rest of the document runs (or not) according to our choices here. 

+ (number) the seed we use to ensure reproducibility
+ (TRUE/FALSE) whether we are running our models on training data
+ (TRUE/FALSE) whether we are running our models on testing data
+ (TRUE/FALSE) run cross-validation on the training set
+ (number) K, the number of CV folds
+ (TRUE/FALSE) process features for training set
+ (TRUE/FALSE) run evaluation on an independent test set
+ (TRUE/FALSE) process features for test set

```{r exp_setup}
set.seed(2018)    # set seed
trainmodel        <- TRUE # if running models on training data
testmodel         <- TRUE  # if running models on testing data
run.cv            <- TRUE  # run cross-validation on the training set
K                 <- 5     # number of CV folds
run.feature.train <- TRUE # process features for training set
run.test          <- TRUE  # run evaluation on an independent test set
run.feature.test  <- TRUE  # process features for test set
```

Using cross-validation or independent test set evaluation, we compare the performance of different classifiers. In this example, we use GBM with different `depth`. In the following code chunk, we list, in a vector, setups (in this case, `depth`) corresponding to model parameters that we will compare. 

????In your project, you will likely be comparing different classifiers than the one considered here and therefore will need to expand this code. You could, for example, assign them numerical IDs and labels specific to your project. 
????At each depth level, makes yes or no response that splits data in half. 

```{r model_setup}
model_values <- seq(5, 10, 2) # trying a few different depths 
model_labels <- paste("GBM with depth =", model_values)
```

## Step 2: Import training and testing images and classification labels.

Below, we code dog as 1 and cat as 0. 
```{r train_label}
if (trainmodel){
label_train <- read.table(paste(experiment_dir, "train_label.txt", sep = ""), header = F)
label_train <- as.numeric(unlist(label_train) == "dog") # turn into a binary -- 1 is dog. 
write.csv(label_train, '../output/label_train.csv')
}

if (testmodel){
label_test <- read.table(paste(experiment_dir, "test_label.txt", sep = ""), header = F)
label_test <- as.numeric(unlist(label_test) == "dog") # turn into a binary -- 1 is dog.
write.csv(label_test, '../output/label_test.csv')
}
```

## Step 3: Construct visual features

????`feature.R` should be the wrapper for all your feature engineering functions and options. The function `feature( )` should have options that correspond to different scenarios for your project and produces an R object that contains features that are required by all the models you are going to evaluate later. 

### SIFT features
```{r SIFT features}
# training
if(run.feature.train){
load('100kmeans.rda')

sift_feature_dir <- "../data/train-features/" 
num_images = length(list.files(path = sift_feature_dir))

data = list()

for (i in 1:num_images){
  file = sprintf("pet%i.jpg.sift.RData", i)
  file_path = paste0(sift_feature_dir,file)
  load(file_path)
  feature_clusters <- get.knnx(k.out$centers, features, 1)$nn.index[,1]
  feature_vec <- rep(0,101)
  feature_vec[1] = i
  for (j in 1:length(feature_clusters)){
  c = feature_clusters[j] + 1 #since the first column is reserved for the index
  feature_vec[c] = feature_vec[c]+ 1}
  data[[i]] = feature_vec
}

sift_data_train = do.call("rbind", data)
}


# testing set
if(run.feature.test){
load('100kmeans.rda')
sift_feature_dir <- "../data/test-features/"
num_images = length(list.files(path = sift_feature_dir))

data = list()

for (i in 1:num_images){  
  file = sprintf("pet%i.jpg.sift.RData", i)
  file_path = paste0(sift_feature_dir,file)
  load(file_path)
  feature_clusters <- get.knnx(k.out$centers, features, 1)$nn.index[,1]
  feature_vec <- rep(0,101)
  feature_vec[1] = i
  for (j in 1:length(feature_clusters)){
  c = feature_clusters[j] + 1 #since the first column is reserved for the index
  feature_vec[c] = feature_vec[c]+ 1}
  data[[i]] = feature_vec
}

sift_data_test = do.call("rbind", data)
}
```

### HOG features
```{r HOG features}

# 
# training data
source("../lib/feature_HOG.R")
tm_feature_train <- NA
if(run.feature.train){
  tm_feature_train <- system.time(hog_train <- feature(img_train_dir, export = TRUE))
  save(hog_train, file = "../output/HOG_features_train.RData")
}
# 
# testing data
source("../lib/feature_HOG_test.R")
tm_feature_test <- NA
if(run.feature.test){
  tm_feature_test <- system.time(hog_test <- feature(img_test_dir, export = TRUE))
  save(hog_test, file = "../output/HOG_features_test.RData")
}

```

```{r}
#Combining Features - SIFTHOG TRAIN
siftfeature_train <- sift_data_train
siftfeature_train$mergeid <- seq(1, nrow(siftfeature_train), 1)

HOGfeature_train <- hog_train
HOGfeature_train <- data.frame(HOGfeature_train[,-1])
HOGfeature_train$mergeid <- seq(1, nrow(HOGfeature_train), 1)
SIFTHOG_train <- merge(HOGfeature_train, siftfeature_train, by=c("mergeid"))
#write.csv(SIFTHOG_train,"../output/SIFTHOG_train.CSV")

#Testing FEATURES
siftfeature_test <- sift_data_test
siftfeature_test$mergeid <- seq(1, nrow(siftfeature_test), 1)

HOGfeature_test <- hog_test
HOGfeature_test <- data.frame(HOGfeature_test[,-1])
HOGfeature_test$mergeid <- seq(1, nrow(HOGfeature_test), 1)
SIFTHOG_test <- merge(HOGfeature_test, siftfeature_test, by=c("mergeid"))
#write.csv(SIFTHOG_test,"../output/SIFTHOG_test.CSV")
```

### Training Baseline Model on Split Training Data

### Step 4: Train a classification model with training images (and the visual features constructed above)

Call the train model and test model from library. 

`train.R` and `test.R` should be wrappers for all your model training steps and your classification/prediction steps. 
+ `train.R`
  + Input: a path that points to the training set features.
  + Input: an R object of training sample labels.
  + Output: an RData file that contains trained classifiers in the forms of R objects: models/settings/links to external trained configurations.
+ `test.R`
  + Input: a path that points to the test set features.
  + Input: an R object that contains a trained classifier.
  + Output: an R object of class label predictions on the test set. If there are multiple classifiers under evaluation, there should be multiple sets of label predictions. 
  
```{r loadlib}
source("../lib/train.R")
source("../lib/test.R")
```
Train function runs GBM function and test just uses the fitted training model and our data to make some predictions. 

#### Model selection with cross-validation

* Do model selection.  Here we choose between model parameters, in this case the interaction depth for GBM. 

```{r runcv, message=FALSE, warning=FALSE}
source("../lib/cross_validation.R")

label.train <- read.csv("../output/label_train.csv",header=TRUE, as.is = TRUE)
label.test <- read.csv("../output/label_test.csv",header=TRUE, as.is = TRUE)


if(run.cv){
  err_cv <- array(dim = c(length(model_values), 2))
  for(k in 1:length(model_values)){
    cat("k=", k, "\n")
    err_cv[k,] <- cv.function(sift_data_train, label.train[,2], model_values[k], K)
  }
  save(err_cv, file = "../output/err_cv.RData")
}
```
Counts number of misclassifications.


This now calculates the best decision. 
```{r cv_vis}
if(run.cv){
  load("../output/err_cv.RData")
  #pdf("../figs/cv_results_baseline.pdf", width=7, height=5)
  plot(model_values, err_cv[,1], xlab = "Interaction Depth", ylab = "CV Error",
       main = "Cross Validation Error", type = "n", ylim = c(0, 0.30))
  points(model_values, err_cv[,1], col = "blue", pch=16)
  lines(model_values, err_cv[,1], col = "blue")
  arrows(model_values, err_cv[,1] - err_cv[,2], model_values, err_cv[,1] + err_cv[,2], 
        length = 0.1, angle = 90, code = 3)
  #dev.off()
}
```

* Choose the "best" parameter value

```{r best_model}
model_best <- model_values[2]
if(run.cv){
  model_best <- model_values[which.min(err_cv[, 1])]
}

par_best <- list(depth = model_best)
```

* Train the model with the entire training set using the selected model (in this case, model parameter) via cross-validation.


```{r final_train}
# need to edit train.R and test.R with our winning model 
tm_train <- NA
tm_train <- system.time(baseline_train <- train(sift_data_train, label.train[ ,2], par_best))
save(baseline_train, file = "../output/baseline_train.RData")
```
## Training Advanced Model

```{r}
label.train <- read.csv("../output/label_train.csv",header=TRUE, as.is = TRUE)
label.test <- read.csv("../output/label_test.csv",header=TRUE, as.is = TRUE)

#Training randomforestmodel
randomforestmodel = randomForest(label.train[,2] ~ ., data = SIFTHOG_train, ntree = 900)

#Training svmmodel
train_svm = function(sift_train){
  fitControl = trainControl(method = 'cv', number = 2)
  svmGrid = expand.grid(C = c(0.5,1,2))
  start_time_svm = Sys.time() 
  svm.fit = train(Label~., 
                  data = sift_train,
                  method = "svmLinear",
                  preProc = c('center', 'scale'),
                  tuneGrid = svmGrid,
                  trControl = fitControl)
  end_time_svm = Sys.time() 
  svm_time = end_time_svm - start_time_svm 
  return(list(fit = svm.fit, time = svm_time))
}

test_svm = function(fit_train, sift_test){
  pred = predict(fit_train$fit, newdata = siftfeature_test)
  return(pred)
}

svm = train_svm(siftfeature_train)


```

## Step 5: Make prediction 

Feed the final training model with the test data.  (Note that for this to truly be 'test' data, it should have had no part of the training procedure used above.) 

```{r test}
#Baseline Model
tm_test <- NA
if(run.test){
  load(file = "../output/baseline_train.RData")
  tm_test <- system.time(pred_test <- test(baseline_train, sift_data_test))
  save(pred_test, file = "../output/pred_test.RData")
}

table(label.test[ ,2], pred_test)
mean(y.test == pred_test) ## 72.8% accuracy rate for baseline with SIFT features and random partitioned training set


```

```{r}
#Advanced Model

#RANDOM FOREST

#load("randomforestSIFT.rda") # called rf

pred_random <- predict(rf, SIFTHOG_test)
pred_random  = as.numeric(pred_random) - 1
error_rate_random <- mean(pred_random != siftfeature_test$Label)

#SVM

load("../output/linearsvmsift.rda") # called svm

test_svm = function(fit_train, sift_test){
  pred = predict(fit_train$fit, newdata = siftfeature_test)
  return(pred)
}

pred_svm = test_svm(svm,SIFTHOG_test)
pred_svm = as.numeric(pred_svm) - 1
error_rate_svm <- 1 - mean(pred_svm == siftfeature_test$Label)

#Logistic Regression

load("../output/logisticmodel.rda")

logistic_pred <- predict(logistic.fit, newdata = siftfeature_test[,2:101], "response")

logistic_pred <- data.frame(logistic_pred)
logistic_pred$V2 = 0.5
for (i in 1:nrow(logistic_pred)){
  logistic_pred[i, 2] <- which.max(logistic_pred[i, ])-1}

pred_lr = logistic_pred
for (i in 1:length(pred_lr)){
  if (pred_lr[i]){ pred_lr[i] = 0
  }
  else{ pred_lr[i] = 1}
}

error_rate_lr = 1 - mean(pred_lr == siftfeature_test$Label)

###
pred_combined = rep(0,length(pred_lr))
for (i in 1:length(pred_combined)){
  vote = pred_random[i] + pred_svm[i] + pred_lr[i]
  if (vote > 1){ pred_combined[i] = 1}
}

error_rate_combined = 1 - mean(pred_combined == siftfeature_test$Label)

```

### Summarize Running Time

Prediction performance matters, so does the running times for constructing features and for training the model, especially when the computation resource is limited. 

Summarizes running time. 
```{r running_time}
cat("Time for constructing training features=", tm_feature_train[1], "s \n")
cat("Time for constructing testing features=", tm_feature_test[1], "s \n") # later
cat("Time for training model=", tm_train[1], "s \n")
cat("Time for making prediction=", tm_test[1], "s \n")
```
